---
title: "Guiding Oncology Dose-Escalation Trials"
author: "Andrew Bean, Sebastian Weber, Lukas Widmer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guiding Oncology Dose-Escalation Trials}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

  
```{r, include=FALSE}
library(OncoBayes2)
library(knitr)
library(ggplot2)
  
theme_set(bayesplot::bayesplot_theme_get())
  
knitr::opts_chunk$set(
  fig.width = 1.62*4,
  fig.height = 4
)
## setup up fast sampling when run on CRAN
is_CRAN <- !identical(Sys.getenv("NOT_CRAN"), "true")
## NOTE: for running this vignette locally, please uncomment the
## following line:
## is_CRAN <- TRUE
.user_mc_options <- list()

if (is_CRAN) {
  .user_mc_options <- options(OncoBayes2.MC.warmup=10, OncoBayes2.MC.iter=20, OncoBayes2.MC.chains=1)
}

options(warn = -1)

```


## Introduction
  
The `OncoBayes2` package provides flexible functions for Bayesian
meta-analytic modeling of the incidence of Dose Limiting Toxicities
(DLTs) by dose level, under treatment regimes involving any number of
combination partners. Such models may be used to ensure patient safety
during trial conduct by supporting dose-escalation decisions. In
addition, the model can support estimation of the Maximum Tolerated
Dose (MTD) in adaptive Bayesian dose-escalation designs.

Whereas traditional dose escalation designs, such as the 3+3 design,
base the dosing decisions on predefined rules about the number of DLTs
in the last one or two cohorts at the current dose, model-based designs
such as those using Bayesian Logistic Regression Models (BLRMs) endeavor
to model the dose-toxicity relationship as a continuous curve, and 
allow the model to guide dosing decisions. In this way, all available data
contributes to the dosing decisions. Furthermore, extensions to the BLRM
approach can support inclusion of available historical data on the 
compound(s) involved.
  
The package supports incorporation of historical data through a
Meta-Analytic-Combined (MAC) framework [1], which stratifies these
heterogeneous sources of information through a hierarchical
model. Additionally, it allows the use of
EXchangeable/Non-EXchangeable (EX/NEX) priors to manage the amount of
information-sharing across subgroups of historical and/or concurrent
sources of data.

```{r, message = FALSE}
## Load involved packages
library(dplyr)  ## for mutate
library(tidyr)  ## defines expand_grid
library(tibble) ## for tibbles
```

## Example use-case

Consider the application described in Section 3.2 of [1], in which the
risk of DLT is to be studied as a function of dose for two drugs, drug
A and drug B. Historical information on the toxicity profiles of these
two drugs is available from single agent trials `trial_A` and
`trial_B`. The historical data for this example is available in an
internal data set.
  
```{r}
kable(hist_combo2)
```

The objective is to aid dosing and dose-escalation decisions in a
future trial, `trial_AB`, in which the drugs will be
combined. Additionally, another investigator-initiated trial `IIT`
will study the same combination concurrently. Note that these
as-yet-unobserved sources of data are included in the input data as
unobserved factor levels. This mechanism allows us to specify a joint
meta-analytic prior for all four sources of historical and concurrent
data.

```{r}
levels(hist_combo2$group_id)
```


### Setting up the trial design

The function `blrm_trial` provides an object-oriented framework for
operationalizing the dose-escalation trial design. This framework is
intended as a convenient wrapper for the main model-fitting engine of
the package, the `blrm_exnex()` function. The latter allows additional
flexibility for specifying the functional form of the model, but
`blrm_trial` covers the most common use-cases. This introductory
vignette highlights `blrm_trial` in lieu of `blrm_exnex`; the reader
is referred to the help-page of the function`?blrm_exnex` for more
details.

One begins with `blrm_trial` by specifying three key design elements:

* The dose-toxicity data
* Information about the study drugs
* The provisional dose levels to be studied during the escalation trial

Information about the study drugs is encoded through a `tibble` as below. 
This includes the names of the study-drugs, the reference doses (see 
[3] or `?blrm_exnex` to understand the role this choice plays in the
model specification), the dosing units, and (optionally) the a priori
expected DLT rate for each study drug given individually at the respective
reference doses.

All design information for the study described in [1] is also included as
built-in datasets, which are part of the `OncoBayes2` package.

#### Drug info

```{r}
kable(drug_info_combo2)
```

#### Dose info

The provisional dose levels are specified as below. For conciseness, we 
begin by fixing the dose level of drug A in these provisional doses.


```{r}
dose_info <- filter(dose_info_combo2, group_id == "trial_AB", drug_A == 3)
kable(dose_info)
```

#### Initializing a `blrm_trial`

Together with the data described in the previous section, these objects
can be used to initialize a `blrm_trial` object.


```{r}
combo2_trial_setup <- blrm_trial(
  data = hist_combo2,
  drug_info = drug_info_combo2,
  dose_info = dose_info,
  simplified_prior = FALSE
)
```



### Specifying the prior and fitting the model

At this point, the trial design has been initialized. However, in the 
absence of `simplified_prior = TRUE`, we have not yet specified the
prior distribution for the dose-toxicity model.

OncoBayes2 provides two methods for completing the model specification:

1. Use `simplified_prior = TRUE`, which employs a package-default
prior distribution, subject to a small number of optional
arguments controlling the details.

2. Provide a full prior specification to be passed to the main model-fitting
engine of OncoBayes2: the `blrm_exnex` function.

For illustrative purposes, here we highlight method #1. See 
`?'example-combo2_trial'` for an example of #2. The below
choice of prior broadly follows the case study in [4], although we 
slightly deviate from the model in [4] by allowing an 
EXchangeable/Non-EXchangeable prior for the drug components.

To employ the simplified prior, and fit the model with MCMC:

```{r, message = FALSE, echo = TRUE, results = "hide"}
combo2_trial_start <- blrm_trial(
  data = hist_combo2,
  drug_info = drug_info_combo2,
  dose_info = dose_info,
  simplified_prior = TRUE,
  EX_prob_comp_hist = 0.8,
  EX_prob_comp_new = 1
)
```

Now, the object `combo2_trial_start` contains the posterior
model fit, in addition to the trial design details. Next we highlight
the main methods for extracting relevant information from it.

### Summary of prior specification

The function `prior_summary` provides a facility for printing, in a
readable format, a summary of the prior specification.

```{r, eval = FALSE}
prior_summary(combo2_trial_start) # not run here
```


### Summary of posterior

The main target of inference is generally the probability of DLT at a
selection of provisional dose levels. To obtain these summaries for
the provisional doses specified previously, we simply write:

```{r}
kable(summary(combo2_trial_start, "dose_prediction"), digits = 2)
```

Such summaries may be used to assess which combination doses have
unacceptable high risk of toxicity. For example, according to the escalation
with overdose control (EWOC) design criteria [3], one would compute the
posterior probability that each dose is excessively toxic (column 
`prob_overdose`; note that the definition of "excessively toxic" is
encoded in the `blrm_trial` object through the `interval_prob` argument),
and take as eligible doses only those where this probability does not
exceed 25% (column `ewoc_ok`).

### Posterior predictive summaries

The BLRM allows a principled approach to predicting the number of DLTs
that may be observed in a future cohort. This may be a key estimand
for understanding and limiting the toxicity risk to patients. For example,
suppose a candidate starting dose for the new trial `trial_AB` is
3 mg of Drug A + 400 mg of Drug B. We may wish to check that at this
dose, the estimated probability of 2 or more DLTs out of
an initial cohort of 3 to 6 patients is sufficiently low. 

```{r}
candidate_starting_dose <- summary(combo2_trial_start, "dose_info") %>%
  filter(drug_A == 3, drug_B == 400) %>%
  expand_grid(num_patients = 3:6) %>%
  mutate(num_toxicities = 0)

pp_summary <- summary(combo2_trial_start, interval_prob = c(-1, 0, 1, 6), predictive = TRUE,
                      newdata = candidate_starting_dose)

kable(bind_cols(select(candidate_starting_dose, num_patients),
                select(pp_summary, ends_with("]"))), digits = 3)
```

This tells us that for the initial cohort, according to
the model, the chance of two or more patients developing DLTs ranges from
`r paste0(100 * round(pp_summary[1, "(1,6]"], 3), "%")` to 
`r paste0(100 * round(pp_summary[4, "(1,6]"], 3), "%")`, 
depending on the number of patients enrolled.

### Updating the model with new data

Dose-escalation designs are adaptive in nature, as dosing decisions are made
after each sequential cohort. The model must be updated after each 
new enrollment. If a new cohort of patients is observed, say:

```{r}
new_cohort <- tibble(group_id = "trial_AB",
                     drug_A = 3,
                     drug_B = 400,
                     num_patients = 5,
                     num_toxicities = 1)
```

One can update the model to incorporate this new information using `update()`
with `add_data` equal to the new cohort:

```{r, message = FALSE, echo = TRUE, results = "hide"}
combo2_trial_update <- update(combo2_trial_start, add_data = new_cohort)
```

This yields a new `blrm_trial` object with updated data and 
posterior summaries. Below is the method for obtaining posterior estimates at 
the next highest dose of drug A.

```{r}
kable(summary(combo2_trial_update, "newdata_prediction",
              newdata = tibble(group_id = "trial_AB",
                               drug_A = 4.5,
                               drug_B = c(400, 600, 800))), digits = 2)
```



### Data scenarios

It may be of interest to test prospectively how this
model responds in various scenarios for upcoming cohorts.

This can be done easily by again using `update()` with the 
`add_data` argument. In the code below, we
explore 3 possible outcomes for a subsequent cohort enrolled at 3 mg
Drug A + 800 mg Drug B, and review the model's inference at adjacent
doses. 

```{r, message = FALSE, echo = TRUE, results = "hide"}
# set up two scenarios at the starting dose level
# store them as data frames in a named list
scenarios <- expand_grid(
  group_id  = "trial_AB",
  drug_A = 3,
  drug_B = 800,
  num_patients = 3,
  num_toxicities = 0:2
) %>% split(1:3) %>% setNames(paste(0:2, "DLTs"))

candidate_doses <- expand_grid(
  group_id = "trial_AB",
  drug_A = c(3, 4.5),
  drug_B = c(600, 800)
)


scenario_inference <- lapply(scenarios, function(scenario_newdata) {
  
  # refit the model with each scenario's additional data
  scenario_fit <- update(combo2_trial_update, add_data = scenario_newdata)
  
  # summarize posterior at candidate doses
  summary(scenario_fit, "newdata_prediction", newdata = candidate_doses)
  
})

```


```{r, echo = FALSE}
kable(scenario_inference[["0 DLTs"]], digits = 2,
      caption = "Model inference when 0 DLTs are observed in the next cohort")
```

```{r, echo = FALSE}
kable(scenario_inference[["1 DLTs"]], digits = 2,
      caption = "Model inference when 1 DLT is observed in the next cohort")
```

```{r, echo = FALSE}
kable(scenario_inference[["2 DLTs"]], digits = 2,
      caption = "Model inference when 2 DLTs are observed in the next cohort")
```

## Continuation of example

In the example of [1], at the time of completion of `trial_AB`, the
following additional data was observed.

```{r}
kable(filter(codata_combo2, cohort_time > 0))
```

Numerous toxicities were observed in the concurrent `IIT`
study. Through the MAC framework, these data can influence the model
summaries for `trial_AB`. Note that we use the `update` function
differently than before, since we specify the entire data-set now we
use the `data` argument.

```{r, message = FALSE, echo = TRUE, results = "hide"}
final_fit <- update(combo2_trial_start, data = codata_combo2)
```


Model results can be visualized using package functions. See 
`?plot_toxicity_curve` for details.

```{r, fig.height = 1.05 * 4, fig.width=1.62 * 4}
plot_toxicity_curve(final_fit, x = vars(drug_B), group = vars(group_id, drug_A),
                    facet_args = list(ncol = 1))

```

## References

[1] Neuenschwander, B., Roychoudhury, S., & Schmidli, H. (2016). On
the use of co-data in clinical trials. Statistics in Biopharmaceutical
Research, 8(3), 345-354.

[2] Neuenschwander, B., Wandel, S., Roychoudhury, S., & Bailey,
S. (2016). Robust exchangeability designs for early phase clinical
trials with multiple strata. Pharmaceutical statistics, 15(2),
123-134.

[3] Neuenschwander, B., Branson, M., & Gsponer, T. (2008). Critical
aspects of the Bayesian approach to phase I cancer trials. Statistics
in medicine, 27(13), 2420-2439.

[4] Neuenschwander, B., Matano, A., Tang, Z., Roychoudhury, S.,
Wandel, S. Bailey, Stuart. (2014). A Bayesian Industry Approach to
Phase I Combination Trials in Oncology. In Statistical methods in drug
combination studies (Vol. 69). CRC Press.

## Session Info

```{r}
sessionInfo()
```


```{r, include=FALSE}
## restore previous global user options
options(.user_mc_options)
```
